---
title: "clickstream"
output: html_document

---


In this analysis we examined the data of Credit Sesame user profile, first session actions and first month engagements.

We clustered the numerical variables via kmeans clustering and combined with categorical variables in logistic regression to predict whether user sign up for a card (click apply) during first session, and visualized different user engagement level. 



# Pre-processing
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Exploratory
```{r}
library(viridis)
library(lubridate)
library(ggraph)
library(tidyverse)
library(sqldf)
#theme
my_theme <- function(base_size = 12, base_family = "Helvetica"){
    theme_minimal() +
    theme(axis.title.y = element_blank(),axis.title.x = element_blank(),
    plot.title = element_text(face="bold", size=16),
    axis.text = element_text(face="bold"),
    plot.background = element_rect(fill = 'ghostwhite',color='white'),
    legend.position = 'None', legend.title = element_blank())
}
```

```{r}
user_profile <- read_csv("~/git_repo/clickstream/dataset/user_profile.csv")
#first_session <- read_csv("~/git_repo/clickstream/dataset/first_session.csv")
#user_engagement <- read_csv("~/git_repo/clickstream/dataset/user_engagement.csv")
```



## kmeans


```{r}
user_numeric = user_profile %>% 
  select(-c(max_cc_utilization_ratio, avg_cc_utilization_ratio,
            tradelines_max_days_since_opened, tradelines_min_days_since_opened,
            zipcode)) %>%
  select_if(is.numeric) 
user = user_numeric%>%
  na.omit()

users <- as.data.frame(scale(user %>% select(-`Unnamed: 0`)))
write.csv(users,'users.csv',row.names = F)
```

```{r}
colSums(is.na(users))
```

select feature -> scale -> find optimal k-> cluster


```{r}
kmean_withinss <- function(k) {
    cluster <- kmeans(users, k)
    return (cluster$tot.withinss)
}

max_k <-16
# Run algorithm over a range of k 
wss <- sapply(2:max_k, kmean_withinss)
plot(wss, type="b")
```


```{r}
user_cluster <- kmeans(users, 4, nstart = 20)
users$cluster = user_cluster$cluster
```

generate complete results with cluster

```{r}
user_id_lookup <- user_profile%>% select(`Unnamed: 0`, user_id)

users_clus <- users %>% select(cluster) %>% 
  cbind(user$`Unnamed: 0`) 
colnames(users_clus) = c('cluster', 'Unnamed')

#user_idx <- user_profile %>% select(c(Unnamed..0, user_id)) %>% cbind(users$cluster)

user_w_cluster <- user_profile %>% 
  inner_join(user_id_lookup) %>% inner_join(users_clus, by =c(`Unnamed: 0`='Unnamed')) %>%
  select(-c(`Unnamed: 0`))
#write.csv(user_w_cluster, 'user_w_cluster.csv')
```


## analyze the clustesr

```{r}
user_long <- users %>% gather (var, value, - cluster) 

user_long %>% 
  ggplot(aes(as.factor(cluster), value)) + geom_boxplot() +
  facet_wrap(~var,scales = 'free') +
  my_theme()
#write.csv(user_long, 'user_long.csv')
```

```{r}
round(prop.table(table(user_w_cluster$age_bucket, user_w_cluster$cluster),2),2)
```

## used unscaled user attributes
```{r}
user_unscaled <- user_w_cluster %>%
  select_if(is.numeric) %>% select(-zipcode) %>%
  gather (var, value, - cluster) 
write.csv(user_unscaled, 'user_unscaled.csv',row.names = F)
```


```{r}
user_smry = user_w_cluster %>% 
  group_by(is_homeowner, age_bucket, credit_score_bucket) %>%
  count()

write.csv(user_smry, 'user_smry.csv')
```




```{r}
user_id_lookup <- user_profile%>% select(Unnamed..0, user_id)
users_clus <- users %>% select(cluster) %>% 
  cbind(user$Unnamed..0) 
colnames(users_clus) = c('cluster', 'Unnamed..0')
#user_idx <- user_profile %>% select(c(Unnamed..0, user_id)) %>% cbind(users$cluster)
user_categorical <- user_profile %>% select_if((is.character)) %>% 
  inner_join(user_id_lookup) %>% inner_join(users_clus)
```


```{r}

```

```{r}
write.csv(user_categorical, 'user_categorical_w_cluster.csv',row.names = F)
```


# first session

```{r}
df_session <- sqldf("
                    SELECT DISTINCT s.user_id,
CASE WHEN c.user_id is not NULL THEN 'Y' ELSE 'N' END as applied
                    from first_session s 
                    LEFT JOIN 
                    (SELECT distinct user_id 
                    from first_session where action_type='CLICK_APPLY') c 
                    ON s.user_id=c.user_id
                    ")
```
```{r}
df_session <- sqldf("
                    SELECT DISTINCT s.user_id,
CASE WHEN c.user_id is not NULL THEN 'Y' ELSE 'N' END as applied
                    from first_session s 
                    LEFT JOIN 
                    (SELECT distinct user_id 
                    from first_session where action_type='CLICK_APPLY') c 
                    ON s.user_id=c.user_id
                    ")
```


```{r}
sort(table(first_session$offer_product))

    # ADVERTISING  PREMIUMSERVICE  AUTO_INSURANCE         PREMIUM   CREDIT_REPAIR       AUTO_LOAN 
    #         163             180             407             568           10500           41797 
    #     BANKING   PERSONAL_LOAN       MORTGAGES  PERSONAL_LOANS CREDIT_PRODUCTS    CREDIT_CARDS 
    #       48062           49230          179196          248069          564602         2545886 
```

```{r}
df_others <- sqldf("
                    SELECT DISTINCT s.user_id,
CASE WHEN c.user_id is not NULL THEN 'Y' ELSE 'N' END as applied
                    from first_session s 
                    LEFT JOIN 
                    (SELECT distinct user_id 
                    from first_session where action_type='CLICK_APPLY'
                    AND offer_product in ('CREDIT_CARDS','CREDIT_PRODUCTS')) c 
                    ON s.user_id=c.user_id
                    WHERE s.offer_product in ('CREDIT_CARDS','CREDIT_PRODUCTS')
                    ")

write.csv(df_others, 'df_others.csv',row.names=F)
```


```{r}
df_loan <- sqldf("
                    SELECT DISTINCT s.user_id,
CASE WHEN c.user_id is not NULL THEN 'Y' ELSE 'N' END as applied
                    from first_session s 
                    LEFT JOIN 
                    (SELECT distinct user_id 
                    from first_session where action_type='CLICK_APPLY'
                    AND offer_product in ('CREDIT_CARDS','CREDIT_PRODUCTS')) c 
                    ON s.user_id=c.user_id
                    WHERE s.offer_product in ('CREDIT_CARDS','CREDIT_PRODUCTS')
                    ")

write.csv(df_credit, 'df_credit.csv',row.names=F)
```



```{r}
df_loans <- sqldf("
                    SELECT DISTINCT s.user_id,
CASE WHEN c.user_id is not NULL THEN 'Y' ELSE 'N' END as applied
                    from first_session s 
                    LEFT JOIN 
                    (SELECT distinct user_id 
                    from first_session where action_type='CLICK_APPLY'
                    AND offer_product in ('PERSONAL_LOANS','MORTGAGES','PERSONAL_LOAN')) c 
                    ON s.user_id=c.user_id
                    WHERE s.offer_product in ('PERSONAL_LOANS','MORTGAGES','PERSONAL_LOAN')
                    ")

write.csv(df_loans, 'df_loans.csv',row.names=F)
```

In order to predict the interested product by user segments and which attributes of these demographics that leads to ultimate sign-up, we looked into which users interact with the offer page of which product and then built a few submodels to inspect what leads to the ultimate sign-up. 
